---
title: "Writing Reports with R Markdown"
author: "Z Lannes"
date: "2022-10-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
library(tidyverse)
library(here)
library(knitr)
```

This report was prepared for the UN. It contains analyses of countries' GDP, life expectancy, and CO2 emissions. Our goal was to determine to what degree a country's economic strength or weakness may be related to its public health status and impact on climate change. We hypothesize that both life expectancy and CO2 emissions increase with a country's GDP.

```{r data, include=FALSE}
#here acts as a function that sets the working directory to the location of the .Rproj file
gapminder_1997 <- read_csv(here("data/gapminder_1997.csv"))
```

```{R gdp_lifexp_1997, echo=FALSE}
#echo=FALSE hides code, but not output
ggplot(gapminder_1997, aes(x=gdpPercap,y=lifeExp,color=continent,size=pop/1000000)) + 
  geom_point() +
  labs(x="GDP Per Capita",y="Life Expectancy",title = "Life Expectancy Based on GDP Per Capita",size = "Population (in millions)")
```

```{r summary_data, include=FALSE}
n_countries <- gapminder_1997 %>%
  select(country) %>%
  n_distinct()

min_GDP <- summarize(gapminder_1997, round(min(gdpPercap)))

max_GDP <- summarize(gapminder_1997,round(max(gdpPercap)))
```
The above plot shows the relationship between GDP per capita and life expectancy for a total of `r n_countries` countries. For this set, economic wealth ranged from minimum of `r min_GDP` to max of `r signif(max_GDP,6)` per capita

| Summary of Data |
|--------|--------|
| Number of countries | `r n_countries`|
| Miniumum GDP per cap | `r min_GDP` |
| Maximum GDP per cap | `r max_GDP` |

```{r tables}
gapminder_1997 %>%
  filter(continent == "Africa") %>%
  kable()
```